Crunching Large Graphs with Commodity Processors

### 2011-04-19 doc/hotpar2011/softxmt-hotpar2011.tex
terrible rework of intro

Many important emerging applications such as social network analysis,
bioinformatics, and sensor networks rely on crunching very large
graphs. Unfortunately, the computational cost of these applications
gets quickly out of hand. For example, in January 2011, Facebook had
500 million active users with an average of 130 friends each
\cite{Facebook:2011p91}. Its friend suggestion algorithm in 2010 ran on
  a 40-node cluster with 72GB memory per node, and, still, new
  suggestions could only be computed every two days
  \cite{Backstrom:2010p90}. Speeding up these applications at a low
  cost would have a large impact in how we deal with large data-sets and make them even more valuable.

@@

The most interesting computational challenge comes from large,
low-diameter, power law graphs: this combination makes extracting
performance difficult. They do not fit in a single
commodity machine's memory. Their low diameter means that they are
difficult to lay out with locality: every vertex needs to be near every
other vertex. Their power law distribution means that a few vertices
generate much more work than the rest, leading to load
imbalances. Together, these properties \todo{including power law distr?} lead to a
difficult conclusion: each edge traversal is likely to lead to a very
expensive inter-node communication.

@@

\item[Low overhead synchronization:] 
%  We must support global synchronization between threads that allows unrelated threads to proceed unblocked.
  Similarly to a regular memory access, long-latency synchronization
  should also cause a thread to yield its core  and allow others to make
  progress. %\todo{repetitive with the previous item}
\end{description}
We discuss our approach to solving each of these challenges in
turn. For each one, we discuss both our current single-node
implementation and ideas for a multi-node implementation.

>>>>

Many important emerging applications such as social network analysis,
bioinformatics, and sensor networks rely on crunching very large
graphs. Unfortunately, the computational cost of these applications
gets quickly out of hand. A Yahoo snapshot of the World Wide Web graph
from 2002 has more than 1.4 billion vertices and 6.6 billion edges
\cite{pegasus, hadi}; the social network Facebook has over 500 million
active users with 130 friends each on average
\cite{Facebook:2011p91}. Applications that process these large
datasets are slow and power-hungry. Speeding up such applications at a
low cost would have a large impact in how we deal with large data-sets
and make them even more valuable.

%For example, in January 2011, Facebook had
%500 million active users with an average of 130 friends each
%\cite{Facebook:2011p91}. Its friend suggestion algorithm in 2010 ran on
%  a 40-node cluster with 72GB memory per node, and, still, new
%  suggestions could only be computed every two days
%  \cite{Backstrom:2010p90}. Speeding up these applications at a low
%  cost would have a large impact in how we deal with large data-sets and make them even more valuable.

@@

The most interesting computational challenge comes from large,
low-diameter, power law graphs: this combination makes extracting
performance difficult. The Web graph mentioned eariler requires 120GB
of storage \cite{pegasus}, difficult to fit in a single commodity machine's
memory. It has an average diameter of 15.6 \cite{hadi}: this low diameter makes it
difficult to lay out with locality, or to partition across a
cluster. Graphs with power law distributions are also difficult to
partition in a balanced way \cite{lang, leskovec}; their few
high-degree vertices also generate much more work than the rest,
leading to load imbalances.

%Together, these properties \todo{including power law distr?} lead to a
%difficult conclusion: each edge traversal is likely to lead to a very
%expensive inter-node communication.

@@

\item[Low overhead synchronization:] 
%  We must support global synchronization between threads that allows unrelated threads to proceed unblocked.
  Similarly to a regular memory access, long-latency synchronization
  should also cause a thread to yield and allow others to make
  progress. %\todo{repetitive with the previous item}
\end{description}
We discuss our approach to solving each of these challenges in
turn. For each one, we discuss both our current single-node
implementation and ideas for a multi-node implementation.
